---
title: "Animated Movie Analysis"
author: "Josh Garzaniti"
date: "2025-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading in Libraries
```{r}
library(tidyr)
library(tidyselect)
library(tidyverse)
library(ggplot2)
library(ggdark)
library(dplyr)
```
##Load in Data

```{r}
Movies = read.csv("G:/My Drive/Personal Projects/movies.csv")

Reviews = read.csv("G:/My Drive/Personal Projects/critic_reviews.csv")

```

##Run summary on initial data

```{r}
summary(Movies)
```

```{r}
summary(Reviews)
```

##Join the two together

```{r}
Movies_with_Reviews = Movies%>%
  left_join(Reviews, by = "movieId")

summary(Movies_with_Reviews)
```
##Clean up the df

```{r}
Movies_with_Reviews = Movies_with_Reviews%>%
  select(-movieId, -movieURL, -creationDate, -criticPageUrl, -isRtUrl, -publicationUrl, -reviewUrl)

Movies_with_Reviews
```

##Changing some data types and updating observations to be the right values

```{r}
str(Movies_with_Reviews)

Movies_with_Reviews = Movies_with_Reviews%>%
  mutate(critic_score = str_extract(critic_score, pattern = "\\d+(?=%)"))%>%
  mutate(audience_score = str_extract(audience_score, pattern = "\\d+(?=%)"))%>%
  mutate(movieYear = as.factor(movieYear))%>%
  mutate(movieRank = as.ordered(movieRank))%>%
  mutate(critic_score = as.numeric(critic_score))%>%
  mutate(audience_score = as.numeric(audience_score))%>%
  mutate(isFresh = ifelse(reviewState == "fresh", 1,0))%>%
  mutate(isRotten = ifelse(reviewState == "rotten", 1, 0))%>%
  mutate(reviewState = as.factor(reviewState))%>%
  mutate(isFresh = as.numeric(isFresh))%>%
  mutate(isRotten = as.numeric(isRotten))%>%
  mutate(isTopCritic = ifelse(isTopCritic == "True", 1, 0))%>%
  mutate(scoreSentiment = as.factor(scoreSentiment))

Movies_with_Reviews
```

##Run some simple correlation tests between review states, scores, and ranks of different movies, years, and critics to see if there's any basic patterns 
Is there a correlation between year and rating?
```{r}
year_audience = cor.test(as.numeric(as.character(Movies_with_Reviews$movieYear)), Movies_with_Reviews$audience_score)

year_audience
  
```
There is a medium positive correlation between year and audience rating that is statistically significant, which means that as animated movies become more modern, audiences tend to give them a higher rating.

Is there a correlation between year and critic ratings then?
```{r}
year_critic = cor.test(as.numeric(as.character(Movies_with_Reviews$movieYear)), Movies_with_Reviews$critic_score)

year_critic
```
It looks like there is a significant but very weak positive correlation between animated movie years and critic scores, so while statistically significant, there's not much of an impact that modern animated movies have on official critic scores.

## Let's look at critics andif there's some critics that are more kind to animated movies than others.

Create a probability variable in terms of "freshness"
```{r}
Movies_with_critics = Movies_with_Reviews%>%
  group_by(criticName)%>%
  summarise(
    total_reviews = n(),
    percent_fresh = mean(reviewState == "fresh")*100)%>%
  arrange(desc(percent_fresh))

Movies_with_critics
```

If we look at this we can see the top 888 rows are critics that all gave completely fresh ratings. This might be a misconception though, because look at how few some of their historical reviews are...some only have 1 or 2 movies reviewed. Let's maybe think about setting a threshold here to filter out critics how haven't reviewed at least 10 movies.






